CreateFiguresReview

### 2025-07-06 manuscript/reviews/ResponseReview1.tex

\begin{quote}
Comments to the Author It was a pleasure to review this manuscript as
someone who has spent years applying survival analysis methods to
psychological data. This is an ambitious manuscript and I believe it
achieves its goals. The key contributions of this manuscript from my
perspective are the creation of custom functions for data wrangling, and
the section on planning studies/simulations for power analysis (of
course, in addition to the introduction to this category of methods and
the practical tutorials provided. I have some comments on this work that
I believe could strengthen this manuscript's contribution.
\end{quote}

\begin{quote}
General comments:
\end{quote}

\begin{quote}
The Introduction is really nicely framed in terms of providing
compelling information for why this class of statistical models would be
useful in some specific areas of psychology (i.e., reaction time data).
Given that there are already a couple of tutorial-types of articles
introducing specific applications of survival analysis in developmental
psychology (references below, both regarding the use of these methods
for behavioural observations of emotional expressions in a multilevel
framework for recurring events), it would be helpful to further frame
the unique contribution that this work adds. The authors may also
consider further emphasizing the specific theoretical contributions that
using these methods may lead to in the areas of reaction time data etc.
\end{quote}

Response: In the revised ms. we frame the unique contribution that this
works adds on page 6: ``\ldots we are not aware of any tutorials that
are aimed specifically at psychological RT (+ accuracy) data, and which
provide worked examples of the key data processing and Bayesian
multilevel regression modelling steps.''

@@

Response: We added a section in the Supplemental Material (section A in
the revised ms.) which explains the types of time-to-event data
typically obtained in RT tasks (detection, discrimination, and bistable
perception tasks). This section makes it clear that our tutorials
pertain only to estimating single-event occurrences. It also visualizes
recurrent events and includes the Lougheed and Stoolmiller tutorials.

\begin{quote}
Specific comments:
\end{quote}

\begin{quote}
The authors use several different terms to refer to time-to-event models
throughout the manuscript. The title refers to EHA, and the running head
refers to Hazard analysis. Survival analysis is also a common term in
Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller \&
Snyder, 2006). Disambiguating the various terms early in the manuscript
and then constantly using one term throughout the manuscript would be
helpful.
\end{quote}

Response: We disambiguate the various terms in the revised ms. on pages
7-8, and we now only use one term throughout.

@@

\begin{quote}
Page 7: The authors state that they are not aware of EHA tutorials for
psychological time to event data. Previoulsy mentioned tutorial articles
by Lougheed and Stoolmiller could be mentioned here--- these are
demonstrating applications to behavioural observation data within
psychology.
\end{quote}

Response: We added these 3 references to the revised ms. on page 6.

\begin{quote}
Page 8: I'm curious about the decision to focus on discrete time rather
than continuous time models. It would help the reader to explain the
decision either way, or refer the reader to Lougheed et al., 2019 who
discuss this distinction and choices that researchers may make regarding
how to handle decisions around time given data formats. I don't work
with reaction time data, but it seems to me that many experiments in
this area would collect data in continuous time intervals, and it's not
clear to me what the benefit of discretizing into time bins is. Now that
I've read the entire manuscript, I see that some information is
presented on this on Page 50, but it would be helpful to make this
information known to the reader earlier in the manuscript.
\end{quote}

Response: We address our choice for discrete-time methods on pages 8-9
of the revised ms.:

``\ldots{} the definition of hazard and the type of models employed
depend on whether one is using continuous or discrete time units. As a
lab, and mainly for practical reasons, we have much more experience
using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given
that RT is typically treated as a continuous variable. However,
continuous forms of EHA require much more data to estimate the
continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991;
Van Zandt, 2000). Thus, by trading a bit of temporal resolution for a
lower number of trials, discrete-time methods seem ideal for dealing
with typical psychological RT data sets for which there are less than
\textasciitilde200 trials per condition per participant (Panis, Schmidt,
et al., 2020). Moreover, as indicated by Allison (2010), learning
discrete-time EHA methods first will help in learning continuous-time
methods, so it seems like a good starting point.''

@@

Response: Thanks for your constructive and encouraging feedback. As we
also noted in our response to the editor, we have made substantial
revisions to the manuscript that include removing repetitive text, and
placing a stronger focus on clarity and brevity in the revised ms.

\begin{quote}
Major comments: 1. Some discussion on the choice of bin width is needed.
Since all of the example time-to-event data is collected in continuous
time, the analyst must discretize it by defining bins. Defining bins is
a key step in the analysis pipeline and so should not be overlooked. It
seems like results---and potentially conclusions drawn---may depend on
the choice of bin width. What factors should the analyst consider when
defining the width of the bins? Does the number of bins matter? Or is
the number of events per bin important? What about other factors? Should
analysts (or potentially these authors too) consider multiple bin widths
when conducting an analysis? An example comparing results when bin width
varies could be both interesting and informative.
\end{quote}

Response: We discuss the issue of defining bins on page 12 of the
revised ms.:

``Third, the width of each time bin will need to be determined. For
instance, in Figure 1B we chose 100 ms in an arbitrary manner. In
reality, however, bin width will need to be set by considering a number
of factors simultaneously. The optimal bin width will depend on (a) the
length of the observation period in each trial, (b) the rarity of event
occurrence, (c) the number of repeated measures (or trials) per
condition per participant, and (d) the shape of the hazard function.
Finding an appropriate bin width in a given user case before fitting
models will require testing a number of options, when calculating and
plotting the descriptive statistics (see section 3.1). The goal is to
find the smallest bin width that is supported by the amount of data
available. Based on our experience, a bin width of 40 ms is a good
starting value when the number of repeated measures is 100 or less. Too
small bin widths will result in erratic hazard functions as many bins
will have no events, and thus hazard estimates of zero. Interestingly,
the time bins do not need to have the same width. For example, Panis
(2020) used larger bins towards the end of the observation period, as
fewer events occurred there.''

@@

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Line 517: What are model weights? Please explain briefly for the
  reader, including how the interpretation differs for loo and waic.
  Also, in the provided reference (Kruz 2023a), I was only able to find
  a discussion of dWAIC, rather than WAIC, so could you please include a
  definition of or reference for dWAIC in your context? Lastly, when
  reporting the model weights in the example in the text, I'd suggest
  including more than 2 digits so that it is clear if the reported
  weights are actually 0 and 1 or if rounding has just made them appear
  so.
\end{enumerate}
\end{quote}

Response: We briefly explain model weights for loo and waic on page 28
of the revised ms. A discussion of WAIC in Kurz (2023a) can be found in
section 4.6.4., so we did not include a reference for dWAIC in the
revised ms. Lastly, we report more digits when reporting the model
weights.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Tutorials 3a and 3b: Rather than simply stating that the frequentist
  models did not converge and resulted in singular fits, it would be
  helpful to provide more discussion. Do you have an idea why the models
  did not converge? Since this is a tutorial paper, more explanation is
  needed. These frequentist examples don't add much to the paper in
  their current form, so I'd either suggest either (a) providing some
  suggestions to the reader of alternative approaches or specific
  strategies that could be used to define a model that will converge or
  (b) removing these examples from the main text and directing the
  reader to the supplement instead.
\end{enumerate}
\end{quote}

Response: We removed the frequentist examples from the main text and
direct the reader to the Tutorials on page 37 of the revised ms. We have
no specific idea why the models did not converge, so we simply mention
on page 37: ``\ldots{} because multilevel generalized linear regression
models often do not converge with complex random-coefficient structures,
we do not discuss them here.''

@@

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{12}
\tightlist
\item
  Line 536: To me, ``subject-specific'' and ``marginal'' mean the
  opposite things. That is, subject-specific is an effect that is
  conditional on an individual while a ``marginal'' effect has been
  marginalized/averaged over individuals. Could you please clarify your
  use of these terms here?
\end{enumerate}
\end{quote}

Response: We erroneously used the term marginal. We removed this term in
the revised ms.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{13}
\tightlist
\item
  Lines 584-587: Since this is a tutorial paper, I'd suggest including
  explanations for why you make changes (a), (b), and (c).
\end{enumerate}
\end{quote}

Response: We now write ``The general process is similar to Tutorial 2a,
except that (a) we use the person-trial data, (b) we use the symmetric
logit link function, and (c) we change the priors (our prior belief is
that conditional accuracy values between 0 and 1 are equally likely).''
on page 33 of the revised ms. The logit is typically used for logistic
regression by experimental psychologists, so we feel that no extra
explanation is needed.

@@

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{14}
\tightlist
\item
  Line 659: What is a ``fully varying effects structure''? Please
  define/explain.
\end{enumerate}
\end{quote}

Response: We explain this on page 24 of the revised ms.: ``We also use
a''keep it maximal'' approach by specifying a full varying (or random)
effects structure. This means that wherever possible we include varying
intercepts and slopes per participant.''

@@

\begin{quote}
Additionally, it may be helpful to use more standard notations such as
h(t\_ij) to indicate the conditional probability that the event occurs
in bin j for individual i, given that the individual had not already
experienced the event in a previous time period (lines 159-160, pg. 9).
\end{quote}

Response: We added indexes for individual and trial in the regression
model equations in section E of the supplemental material. Furthermore,
our new description on pages 8 and 9, together with the new section A in
the supplemental material, should make this issue clear for the reader.

\begin{quote}
Similarly, the survival probability is the probability that the event
occurs after bin j, \emph{given that the individual lhas survived up
until bin j}.
\end{quote}

Response: We define S(t) = P(RT \textgreater{} t) now as ``the
probability that the event does not occur before the endpoint of bin t''
in section B of the Supplemental Material.

@@

Response: We address our choice for discrete-time methods on pages 8-9
of the revised ms.:

``\ldots{} the definition of hazard and the type of models employed
depend on whether one is using continuous or discrete time units. As a
lab, and mainly for practical reasons, we have much more experience
using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given
that RT is typically treated as a continuous variable. However,
continuous forms of EHA require much more data to estimate the
continuous-time hazard (rate) function well (Bloxom, 1984; Luce, 1991;
Van Zandt, 2000). Thus, by trading a bit of temporal resolution for a
lower number of trials, discrete-time methods seem ideal for dealing
with typical psychological RT data sets for which there are less than
\textasciitilde200 trials per condition per participant (Panis, Schmidt,
et al., 2020). Moreover, as indicated by Allison (2010), learning
discrete-time EHA methods first will help in learning continuous-time
methods, so it seems like a good starting point.''

\begin{quote}
Clearer description of the data would be helpful for readers (as well as
only having this description in one place in the tutorial to be more
concise). For example, on pg. 10 it is indicated that 200 trials of 1
experimental condition are simulated. Can the event occur at any time in
those 200 trials? If the event occurs in trial 150, do the remaining 50
trials still occur, or does the time series end for that person? How are
those extra trials after an event occurred handled in the data wrangling
and analysis. Can events occur more than once?
\end{quote}

Response: We think that there is a misunderstanding of what ``trial''
means. We now provide a clear description of the meaning of ``trial''
and of the various types of time-to-data obtained in typical RT tasks in
section A of the Supplemental Material.

@@

\begin{quote}
The numbered list on pg. 11 contained jargon statements without
explanations. Suggest to simplify and keep the focus on the data you are
using and specific research questions you are using survival analysis as
a tool to address.
\end{quote}

Response: We removed the numbered list in the revised ms.

\begin{quote}
In section 2.3, suggest to lead with the data and research questions of
interest and then show how features of survival analysis make it a good
method to examine the research questions. I also felt that up until now,
the tutorial has been very confusing and hard to follow. It could be
that starting the tutorial at section 2.3 (with a few extra details
filled in), may be an alternative.
\end{quote}

Response: Section 2.3 has been rewritten. The new Figure 1 in the
tutorial directly focuses on the comparison between analysing means
vs.~distributional shapes. We also discuss research questions on page 10
of the revised ms.

@@

\begin{quote}
Section 4.1.2 was a clearer data description than had been presented up
until that point in the tutorial. Again, it is suggested that the
manuscript/tutorial could be dramatically reduced in length if
information was more concisely but descriptively written in a single
place (e.g., data descriptions).
\end{quote}

Response: We provide the information on data descriptions in a single
place in the revised ms.

\begin{quote}
Suggest to be more precise with the language on pg. 21 (line 355) --
which data are nested within participants? The trial data? The binned
data?
\end{quote}

Response: We specify ``person-trial data'' in the revised ms.

@@

\begin{quote}
pg. 22 -- since this is a tutorial, it might also be helpful to unpack
in more detail the section about identifying individuals who may be
guessing etc -- especially since these are very important
considerations.
\end{quote}

Response: We know of no more detail to report.

\begin{quote}
section 4.3.1 -- (line 451) it would be helpful if more details is given
on how to set the analysis time window. What counts as ``enough data''?
Given that this is a tutorial, these types of detail are critical for
readers to be able to implement the method in their own work
\end{quote}

Response: We changed this sentence into: ``\ldots{} because the first
few bins typically contain no responses, one has to select an analysis
time window, i.e., a contiguous set of bins for which there is data for
each participant.'' on page 24 of the revised ms.

@@

\begin{quote}
section 4.3.1 (line 453) -- more detail is needed on the cloglog link
function and how it differs from the more commonly known about logit
link.
\end{quote}

Response: A visual comparison between the cloglog and logit link
functions is provided in section C of the Supplemental Material.

\begin{quote}
section 4.3.1 (line 456) -- what goes into the decision about whether to
treat time as a categorical or continuous predictor?
\end{quote}

Response: We added the following information on page 24 of the revised
ms.: ``Third, one has to choose whether to treat TIME (i.e., the time
bin index t) as a categorical or continuous predictor (see also section
E of the Supplemental Material). For example, when you want to know if
cloglog-hazard is changing linearly or quadracticly over time, you
should treat TIME as a continuous predictor. When you are only
interested in the effect of covariates on hazard, you can treat TIME as
a categorical predictor (i.e., fit an intercept for each bin), in which
case you can choose between reference coding and index coding.''

@@

\begin{quote}
section 4.3.3. -- it might be helpful to give a brief indicator up front
to readers about the rationale/purpose of fitting both M0i and M1i,
especially given that no interpretation is given for M0i. Similarly, it
would be helpful to give a bit more detail on the purpose of the
comparison of M0i and M1i (given that this is a didactic tutorial).
\end{quote}

Response: We mention now that model M0i is a baseline or reference
model, to be compared with model M01 (see page 26 of the revised ms.).

\begin{quote}
section 4.3.6. -- same thing here -- since this is a tutorial, at least
brief details on what the purpose of this section is would be helpful.
It might also be helpful here to define what is meant by credible
interval so that more novice readers don't confuse it with a confidence
interval.
\end{quote}

Response: We added the purpose (``To make causal inferences\ldots{}'')
and define a credible interval on page 29 of the revised ms.
